### 2012-05-31 docs/SDC/sdc.tex
some labels
changed "predict_rhoprime_and_rhoX" to "predict_rhoX".  After we predict rhoX to edges, we set rho = sum(rhoX).  Also, in the enthalpy prediction types 2 and 4, we assemble (rho h) on edges using h_edge * sum(rho X_k)_edge instead of h_edge * (rho_0 + rho'_edge)
final updates for the day
saving progress - new alg notes

@@

\section{SDC Without Thermal Diffusion or Base State Evolution}
In the SDC version, the {\tt VODE} integration at the end of an SDC
iteration is responsible for updating all the thermodynamic quantities
including both the advection (incorporated via a piecewise constant advective 
flux divergence source term) and the reactions.  This provides a much stronger coupling between
the physical processes.  In particular, our system now looks like:
\begin{eqnarray}
\frac{d(\rho X_k)}{dt} &=& \rho \omegadot_k(\rho,X_k,T) + \underbrace{\nabla\cdot(\rho X_k\Ub)}_{A_{\rho X_k}}\label{eq:sdc:rhoX} \\
\frac{d(\rho h)}{dt}   &=& \rho \Hnuc + \underbrace{\nabla\cdot(\rho h\Ub)}_{A_{\rho h}} \label{eq:sdc:rhoh}
\end{eqnarray}
Here, $A_{\rho X_k}$ and $A_{\rho h}$ are piecewise-constant (in time)
approximations to the change in ${\rho X_k}$ and ${\rho h}$ (respectively)
due to the advection.  These are constructed by calling {\tt density\_advance}
and {\tt enthalpy\_advance} in \maestro\ and passed into the network solver
during the reaction step.  A flowchart of the \maestro\ SDC algorithm is 
shown in figure~\ref{fig:sdc:flowchart}.

@@

{\bf Step 1:} ({\it Predictor}) In this step, we integrate $(\rho, \rho X_k, \rho h)$ over the full time step.  The quantities $(S, \beta_0, k_{\rm th}, c_p, \xi_k, h_p)^n$ are computed from the the thermodynamic variables at $t^n$.  This step is divided into several sub-steps:\\

{\bf Step 1A:} ({\it Compute advection velocities}) Use $\Ub^n$ and a second-order Godunov method to compute time-centered edge velocities, $\uadvsdcstar$, with time-lagged dynamic pressure and explicit buoyancy as forcing terms.  The $\star$ superscript indicates that this field does not satisfy the divergence constraint.  Compute $S^{n+\myhalf,{\rm pred}}$ by extrapolating in time,
\begin{equation}
S^{n+\myhalf,{\rm pred}} = S^n + \frac{\Delta t^n}{2}\frac{S^n - S^{n-1}}{\Delta t^{n-1}},
\end{equation}
and project $\uadvsdcstar$ to obtain $\uadvsdcpred$, which satisfies
\begin{equation}
\nabla\cdot\left(\beta_0^n\uadvsdcpred\right) = S^{n+\myhalf,\rm{pred}}.
\end{equation}

{\bf Step 1B:} ({\it Compute advective fluxes}) Use $\uadvsdcpred$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf,{\rm pred}}$, with time-lagged reactions, explicit diffusion, and time-centered thermodynamic pressure as source terms.\\

{\bf Step 1C:} ({\it Compute diffusive fluxes}) Begin by compute a preliminary time-advanced density using
\begin{equation}
\frac{\rho^{n+1,{\rm pred}} - \rho^n}{\Delta t} = -\sum_m\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf,{\rm pred}}\uadvsdcpred\right].
\end{equation}
Then, compute preliminary, time-advanced species using
\begin{equation}
\frac{\rho^{n+1,{\rm pred}}X_{k,{\rm AD}}^{n+1,{\rm pred}} - (\rho X_k)^n}{\Delta t} = -\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf,{\rm pred}}\uadvsdcpred\right] + I_{\rho X_k}^{\rm lagged}.\label{eq:sdc species 2}
\end{equation}
Then, solve a Crank-Nicolson-type diffusion equation for $h^{n+1,{\rm pred}}_{\rm AD}$, using transport coefficients evaluated at $t^n$ everywhere,
\begin{eqnarray}
\frac{\rho^{n+1,{\rm pred}}h^{n+1,{\rm pred}}_{\rm AD} - (\rho h)^n}{\Delta t} &=& -\nabla\cdot\left[\left(\rho h\right)^{n+\myhalf,{\rm pred}}\uadvsdcpred\right] + \uadvsdcpred\cdot\nabla p_0 + I_{\rho h}^{\rm lagged}\nonumber\\
&& + \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n + \nabla\cdot\frac{\kth^n}{c_p^n}\nabla h_{\rm AD}^{n+1,{\rm pred}}\right)\nonumber\\
&& + \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_{k,{\rm AD}}^{n+1,{\rm pred}}\right)\nonumber\\
&& - \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0\right).\label{eq:sdc enthalpy 2}
\end{eqnarray}

{\bf Step 1D:} ({\it Advance thermodynamic variables}) Define $Q_{\rho X_k}^{\rm pred}$ as the right hand side of (\ref{eq:sdc species 2}) without the $I_{\rho X_k}^{\rm lagged}$ term, and define $Q_{\rho h}^{\rm pred}$ as the right hand side of (\ref{eq:sdc enthalpy 2}) without the $I_{\rho h}^{\rm lagged}$ term.  Use {\tt VODE} to integerate (\ref{eq:sdc species}) and (\ref{eq:sdc enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1,{\rm pred}}$ using the piecewise-constant advection and diffusion source terms:
\begin{eqnarray}
\frac{\partial(\rho X_k)}{\partial t} &=& Q_{\rho X_k}^{\rm pred} + \rho\dot\omega_k\\
\frac{\partial(\rho h)}{\partial t} &=& Q_{\rho h}^{\rm pred} + \rho\Hnuc.
\end{eqnarray}
At this point we can define $I_{\rho X_k}^{\rm pred}$ and $I_{\rho h}^{\rm pred}$, or whatever term we need depending on our species and enthalpy edge state prediction types, for use in the corrector step.\\

{\bf Step 2:} ({\it Corrector}) In the corrector, we use the time-advanced state from the predictor to perform a more accurate integration of the thermodynamic variables.  The quantities $(S, \beta_0, k_{\rm th}, c_p, \xi_k, h_p)^{n+1,{\rm pred}}$ are computed from $(\rho,\rho X_k,\rho h)^{n+1,{\rm pred}}$.  This step is divided into several sub-steps:\\\\

{\bf Step 2A:} ({\it Compute advection velocities}) First, compute $S^{n+\myhalf}$ and $\beta_0^{n+\myhalf}$ using
\begin{equation}
S^{n+\myhalf} = \frac{S^n + S^{n+1,{\rm pred}}}{2}, \qquad \beta_0^{n+\myhalf} = \frac{\beta_0^n + \beta_0^{n+1,{\rm pred}}}{2}.
\end{equation}
Then, project $\uadvsdcstar$ to obtain $\uadvsdc$, which satisfies
\begin{equation}
\nabla\cdot\left(\beta_0^{n+\myhalf}\uadvsdc\right) = S^{n+\myhalf}.
\end{equation}

{\bf Step 2B:} ({\it Compute advective fluxes}) Use $\uadvsdc$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf}$, with iteratively-lagged reactions, explicit diffusion, and time-centered thermodynamic pressure as source terms.\\

{\bf Step 2C:} ({\it Compute diffusive fluxes}) Begin by computing the time-advanced density using
\begin{equation}
\frac{\rho^{n+1} - \rho^n}{\Delta t} = -\sum_m\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf}\uadvsdc\right].
\end{equation}
Then, compute preliminary, time-advanced species using
\begin{equation}
\frac{\rho^{n+1}X_{k,{\rm AD}}^{n+1} - (\rho X_k)^n}{\Delta t} = -\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf}\uadvsdc\right] + I_{\rho X_k}^{\rm pred}.\label{eq:sdc species 3}
\end{equation}
Then, solve a backward-Euler-type correction equation for $h^{n+1}_{\rm AD}$,
\begin{eqnarray}
\frac{\rho^{n+1}h^{n+1}_{\rm AD} - (\rho h)^n}{\Delta t} &=& -\nabla\cdot\left[\left(\rho h\right)^{n+\myhalf}\uadvsdc\right] + \uadvsdc\cdot\nabla p_0 + I_{\rho h}^{\rm pred}\nonumber\\
&& + \nabla\cdot\frac{\kth^{n+1,{\rm pred}}}{c_p^{n+1,{\rm pred}}}\nabla h^{n+1}_{\rm AD} + \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n - \nabla\cdot\frac{\kth^{n+1,{\rm pred}}}{c_p^{n+1,{\rm pred}}}\nabla h^{n+1,{\rm pred}}\right)\nonumber\\
&& + \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^{n+1,{\rm pred}} k_{\rm th}^{n+1,{\rm pred}}}{c_p^{n+1,{\rm pred}}}\nabla X_{k,{\rm AD}}^{n+1}\right)\nonumber\\
&& - \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^{n+1,{\rm pred}}k_{\rm th}^{n+1,{\rm pred}}}{c_p^{n+1,{\rm pred}}}\nabla p_0\right)\label{eq:sdc enthalpy 3}
\end{eqnarray}

{\bf Step 2D:} ({\it Advance thermodynamic variables}) Define $Q_{\rho X_k}$ as the right hand side of (\ref{eq:sdc species 3}) without the $I_{\rho X_k}^{\rm pred}$ term, and define $Q_{\rho h}$ as the right hand side of (\ref{eq:sdc enthalpy 3}) without the $I_{\rho h}^{\rm pred}$ term.  Use {\tt VODE} to integerate (\ref{eq:sdc species}) and (\ref{eq:sdc enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1}$ using the piecewise-constant advection and diffusion source terms:
\begin{eqnarray}
\frac{\partial(\rho X_k)}{\partial t} &=& Q_{\rho X_k} + \rho\dot\omega_k\\
\frac{\partial(\rho h)}{\partial t} &=& Q_{\rho h} + \rho\Hnuc.
\end{eqnarray}
At this point we can define $I_{\rho X_k}^{\rm lagged}$ and $I_{\rho h}^{\rm lagged}$, or whatever term we need depending on our species and enthalpy edge state prediction types, for use in the predictor in the next time step.\\

{\bf Step 3:} ({\it Advance velocity and dynamic pressure}) Same as the original \maestro\ algorithm.

>>>>

\label{ch:sdc}

@@

\section{SDC Without Thermal Diffusion or Base State Evolution}
In the SDC version, the {\tt VODE} integration at the end of an SDC
iteration is responsible for updating all the thermodynamic quantities
including both the advection (incorporated via a piecewise constant advective 
flux divergence source term) and the reactions.  This provides a much stronger coupling between
the physical processes.  In particular, our system now looks like:
\begin{eqnarray}
\frac{d(\rho X_k)}{dt} &=& \rho \omegadot_k(\rho,X_k,T) + \underbrace{\nabla\cdot(\rho X_k\Ub)}_{A_{\rho X_k}}\label{eq:sdc:rhoX} \\
\frac{d(\rho h)}{dt}   &=& \rho \Hnuc + \underbrace{\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt}}_{A_{\rho h}} \label{eq:sdc:rhoh}
\end{eqnarray}
Here, $A_{\rho X_k}$ and $A_{\rho h}$ are piecewise-constant (in time)
approximations to the change in ${\rho X_k}$ and ${\rho h}$ (respectively)
due to the advection.  These are constructed by calling {\tt density\_advance}
and {\tt enthalpy\_advance} in \maestro\ and passed into the network solver
during the reaction step.  A flowchart of the \maestro\ SDC algorithm is 
shown in figure~\ref{fig:sdc:flowchart}.

@@

{\bf Step 1:} ({\it Compute advection velocities})\\ \\
Use $\Ub^n$ and a second-order Godunov method to compute time-centered edge velocities, $\uadvsdcstar$, with time-lagged dynamic pressure and explicit buoyancy as forcing terms.  The $\star$ superscript indicates that this field does not satisfy the divergence constraint.  Compute $S^{n+\myhalf,{\rm pred}}$ by extrapolating in time,
\begin{equation}
S^{n+\myhalf,{\rm pred}} = S^n + \frac{\Delta t^n}{2}\frac{S^n - S^{n-1}}{\Delta t^{n-1}},
\end{equation}
and project $\uadvsdcstar$ to obtain $\uadvsdcpred$, which satisfies
\begin{equation}
\nabla\cdot\left(\beta_0^n\uadvsdcpred\right) = S^{n+\myhalf,\rm{pred}}.
\end{equation}

{\bf Step 2:} ({\it Predictor})\\ \\
 In this step, we integrate $(\rho, \rho X_k, \rho h)$ over the full time step.  The quantities $(S, \beta_0, k_{\rm th}, c_p, \xi_k, h_p)^n$ are computed from the the thermodynamic variables at $t^n$.  This step is divided into several sub-steps:\\

{\bf Step 2A:} ({\it Compute advective flux divergences})\\ \\
Use $\uadvsdcpred$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf,(0)}$, with time-lagged reactions ($I^{\rm lagged} = I^{(j_{\rm max})}$ from the previous time step), explicit diffusion, and time-centered thermodynamic pressure as source terms.  Define the advective flux divergences as
\begin{eqnarray}
A_{\rho X_k}^{(0)} &=& -\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf,{(0)}}\uadvsdcpred\right],\\
A_{\rho h}^{(0)} &=& -\nabla\cdot\left[\left(\rho h\right)^{n+\myhalf,(0)}\uadvsdcpred\right].
\end{eqnarray}
Next, use these fluxes to compute a preliminary time-advanced density using
\begin{equation}
\frac{\rho^{n+1,(0)} - \rho^n}{\Delta t} = \sum_mA_{\rho X_k}^{(0)}.
\end{equation}
Then, compute preliminary, time-advanced species using
\begin{equation}
\frac{\rho^{n+1,(0)}X_{k,{\rm AD}}^{n+1,(0)} - (\rho X_k)^n}{\Delta t} = A_{\rho X_k}^{(0)} + I_{\rho X_k}^{\rm lagged}.\label{eq:sdc species 2}
\end{equation}

{\bf Step 2B:} ({\it Compute diffusive flux divergence})\\ \\
Solve a Crank-Nicolson-type diffusion equation for $h^{n+1,(0)}_{\rm AD}$, using transport coefficients evaluated at $t^n$ everywhere,
\begin{eqnarray}
\frac{\rho^{n+1,(0)}h^{n+1,(0)}_{\rm AD} - (\rho h)^n}{\Delta t} &=& A_{\rho h}^{(0)} + \uadvsdcpred\cdot\nabla p_0 + I_{\rho h}^{\rm lagged}\nonumber\\
&& + \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n + \nabla\cdot\frac{\kth^n}{c_p^n}\nabla h_{\rm AD}^{n+1,(0)}\right)\nonumber\\
&& + \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_{k,{\rm AD}}^{n+1,(0)}\right)\nonumber\\
&& - \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0\right).\label{eq:sdc enthalpy 2}
\end{eqnarray}

{\bf Step 2C:} ({\it Advance thermodynamic variables})\\ \\
Define $Q_{\rho X_k}^{(0)}$ as the right hand side of (\ref{eq:sdc species 2}) without the $I_{\rho X_k}^{\rm lagged}$ term, and define $Q_{\rho h}^{(0)}$ as the right hand side of (\ref{eq:sdc enthalpy 2}) without the $I_{\rho h}^{\rm lagged}$ term.  Use {\tt VODE} to integerate (\ref{eq:sdc species}) and (\ref{eq:sdc enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1,(0)}$ using the piecewise-constant advection and diffusion source terms:
\begin{eqnarray}
\frac{\partial(\rho X_k)}{\partial t} &=& Q_{\rho X_k}^{(0)} + \rho\dot\omega_k\\
\frac{\partial(\rho h)}{\partial t} &=& Q_{\rho h}^{(0)} + \rho\Hnuc.
\end{eqnarray}
At this point we can define $I_{\rho X_k}^{(0)}$ and $I_{\rho h}^{(0)}$, or whatever term we need depending on our species and enthalpy edge state prediction types, for use in the corrector step.\\

{\bf Step 3:} ({\it Update advection velocities})\\ \\
First, compute $S^{n+\myhalf}$ and $\beta_0^{n+\myhalf}$ using
\begin{equation}
S^{n+\myhalf} = \frac{S^n + S^{n+1,(0)}}{2}, \qquad \beta_0^{n+\myhalf} = \frac{\beta_0^n + \beta_0^{n+1,(0)}}{2}.
\end{equation}
Then, project $\uadvsdcstar$ to obtain $\uadvsdc$, which satisfies
\begin{equation}
\nabla\cdot\left(\beta_0^{n+\myhalf}\uadvsdc\right) = S^{n+\myhalf}.
\end{equation}

{\bf Step 4:} ({\it Corrector Loop})\\ \\
We loop over this step from $j=1,j_{\rm max}$ times.  In the corrector, we use the time-advanced state from the predictor to perform a more accurate integration of the thermodynamic variables.  The quantities $(S, \beta_0, k_{\rm th}, c_p, \xi_k, h_p)^{n+1,(j-1)}$ are computed from $(\rho,\rho X_k,\rho h)^{n+1,(j-1)}$.  This step is divided into several sub-steps:\\

{\bf Step 4A:} ({\it Compute advective flux divergences})\\ \\
Use $\uadvsdc$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf}$, with iteratively-lagged reactions ($I^{(j-1)}$), explicit diffusion, and time-centered thermodynamic pressure as source terms.  Define the advective flux divergences as
\begin{eqnarray}
A_{\rho X_k}^{(j)} &=& -\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf,(j)}\uadvsdc\right],\\
A_{\rho h}^{(j)} &=& -\nabla\cdot\left[\left(\rho h\right)^{n+\myhalf,(j)}\uadvsdc\right].
\end{eqnarray}
Next, use these fluxes to compute the final time-advanced density using
\begin{equation}
\frac{\rho^{n+1} - \rho^n}{\Delta t} = -\sum_mA_{\rho X_k}^{(j)}.
\end{equation}
Then, compute preliminary, time-advanced species using
\begin{equation}
\frac{\rho^{n+1}X_{k,{\rm AD}}^{n+1,(j)} - (\rho X_k)^n}{\Delta t} = A_{\rho X_k}^{(j)} + I_{\rho X_k}^{(j-1)}.\label{eq:sdc species 3}
\end{equation}

{\bf Step 4B:} ({\it Compute diffusive flux divergence})\\ \\
Solve a backward-Euler-type correction equation for $h^{n+1,(j)}_{\rm AD}$,
\begin{eqnarray}
\frac{\rho^{n+1}h^{n+1,(j)}_{\rm AD} - (\rho h)^n}{\Delta t} &=& A_{\rho h}^{(j)} + \uadvsdc\cdot\nabla p_0 + I_{\rho h}^{(j-1)}\nonumber\\
&& + \nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla h^{n+1,(j)}_{\rm AD} + \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n - \nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla h^{n+1,(j-1)}\right)\nonumber\\
&& + \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^{n+1,(j-1)} k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla X_{k,{\rm AD}}^{n+1,(j)}\right)\nonumber\\
&& - \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^{n+1,(j-1)}k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla p_0\right)\label{eq:sdc enthalpy 3}
\end{eqnarray}

{\bf Step 4C:} ({\it Advance thermodynamic variables})\\ \\
Define $Q_{\rho X_k}^{(j)}$ as the right hand side of (\ref{eq:sdc species 3}) without the $I_{\rho X_k}^{(j-1)}$ term, and define $Q_{\rho h}^{(j)}$ as the right hand side of (\ref{eq:sdc enthalpy 3}) without the $I_{\rho h}^{(j-1)}$ term.  Use {\tt VODE} to integerate (\ref{eq:sdc species}) and (\ref{eq:sdc enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1}$ using the piecewise-constant advection and diffusion source terms:
\begin{eqnarray}
\frac{\partial(\rho X_k)}{\partial t} &=& Q_{\rho X_k}^{(j)} + \rho\dot\omega_k\\
\frac{\partial(\rho h)}{\partial t} &=& Q_{\rho h}^{(j)} + \rho\Hnuc.
\end{eqnarray}
At this point we can define $I_{\rho X_k}^{(j)}$, $I_{\rho h}^{(j)}$, and any other $I$ terms we need depending on 
our species and enthalpy edge state prediction types, for use in the predictor in the next time step.\\

{\bf Step 5:} ({\it Advance velocity and dynamic pressure})\\ \\
Similar to the original \maestro\ algorithm, more to come.
